---
output:
  html_document: default
  pdf_document: default
---
Explore red wine dataset
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

#install.packages("ggplot2")
library(ggplot2)
#install.packages("tidyr")
library(tidyr)
library(dplyr)
#install.packages("gridExtra")
library(gridExtra)
#install.packages("devtools")
library(devtools)
#install_github("ggobi/ggally")
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
df <- read.csv("wineQualityReds.csv")
head(df)
```

> I have decided to explore red wine dataset. This tidy dataset contains 1,599 red wines with 11 variables on the chemical properties of the wine. At least 3 wine experts rated the quality of each wine, providing a rating between 0 (very bad) and 10 (very excellent).

## Univariate Plots Section

```{r echo=FALSE}
#remote X as we do not need it
df <- df[,2:length(df)]
summary(df)
```

> Dataset contains 11 quantitative columns (all except quality) and 1 qualitative column (quality)

```{r echo=FALSE}
# add quality factor 
df$quality.factor = factor(df$quality)
str(df)
```

I have created quality.factor which is factor of "quality" this makes sence since it's categorical variable

```{r echo=FALSE, "All columns histogram"}
ggplot(gather(df[,1:12]), aes(value)) +
  ggtitle("Histogram for each column") +
  geom_histogram(bins = 20) +
  facet_wrap(~key, scales = "free")
```

>By looking at data I see that most of the columns might have some outliers, residual.sugar, 
chlorides and sulfates are most visible.
Alcohol, free sulfur.dioxide and total.sulfur.dioxide are clear right-skewed distribution 
while density and ph seem to be normally distributed.
Most of the red wines are quality 5 or 6


```{r echo=FALSE}
ggplot(df, aes(x = citric.acid)) +
  ggtitle("Histogram of Citric acid") +
  geom_histogram()
 
```

After having a closer look at citric.acid, I think it may be a good idea to split values into groups 
as follow 0-0.20, 0.21-0.60 and above 0.6. Distributions seem to be right skewed

```{r echo=FALSE}
df <- df %>% mutate(citric_acid_grouped = ifelse(citric.acid <= 0.2, "0-0.2", 
                                          ifelse(citric.acid > 0.2 & 
                                                   citric.acid < 0.6, 
                                                 "0.2-0.6", "Above 0.6")))

```

```{r echo=FALSE}
ggplot(df, aes(x = chlorides)) +
  ggtitle("Histogram of chlorides") +
  geom_histogram()
 
```

Chlorides seem to be concetrated around 0.1 mark. Threre outliers around 0.4 and 0.6


```{r echo=FALSE}
ggplot(df, aes(x = residual.sugar)) +
  ggtitle("Histogram of Residual sugar") +
  geom_histogram()
 
```

Residual sugar seem to be concentrated around 2 mark. There some ouliers from 8.


```{r echo=FALSE}
ggplot(df, aes(alcohol)) + 
        geom_histogram()
```

```{r echo=FALSE}
# changed binwidth to better see bars 
ggplot(df, aes(alcohol, fill = 'b')) + 
        geom_histogram(binwidth = .01) +
        scale_x_continuous(breaks = seq(7, 15, .5)) + 
        guides(fill = F)
 
```

```{r echo=FALSE}
# checking the quartile values 
summary(df$alcohol)

# creating a new variable 
df$alcohol_content = cut(df$alcohol, breaks = c(0, 9.5, 11.1, 15))

levels(df$alcohol_content) = c("(0,9.5]" = "Low", 
                                 "(9.5,11.1]" = "Medium", 
                                 "(11.1,15]" = "High")

# creating graph to display new variable
ggplot(df, aes(alcohol_content)) + 
        geom_bar()
 
```

Created a new variable (alcohol_content) with low, medium, high alcohol content. There is a strict value (x=9.5%) where alcohol content is optimally chosen for wine. The count gradually decreases as content increase and immediately drops as content decreases below 9.5%. The average alcohol content is 10.42%.


# Univariate Analysis

### What is the structure of your dataset?
>There are 1599 redwine observations in the dataset with 12 features. All 
features contains numeric data except the quality which is a categorical.

### What is/are the main feature(s) of interest in your dataset?
> The main feature is quality 

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

>The other feature that may support the investigation, in my opinion, 
are alcohol, density, ph

### Did you create any new variables from existing variables in the dataset?
> Yes, I have added "citric_acid_grouped" and I have removed "X". Then I created alcohol_content which consists of 3 buckets (0, 9.5], (9.5, 11.1], and (11.1, 15]. Then I renamed each bucket to Low, Medium, and High to differentiate the amount of alcohol catagorically. Then I have added quality.factor to make use of category when drawing graphs

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?
>I have noticed that citric.acid seem to be right skewed and it would fit better if in groups



# Bivariate Plots Section

```{r echo=FALSE}

library(GGally)
ggscatmat(df, columns = 1:12) + theme_grey(base_size=6)
```

The four highest correlation coefficients with quality are:

1. alcohol at 0.48
2. sulphates at 0.25
3. citric.acid at 0.23
4. fixed.acidity at 0.12


```{r echo=FALSE}
ggplot(df, aes(density, alcohol)) + 
  geom_jitter(alpha=0.2) + 
  stat_smooth(method = "lm", formula = y ~ x)

```

At the high and lowest points of alcohol, there is not much density. But there is a trend towards higher density as alcohol content drops.


```{r echo=FALSE}
ggplot(df, aes(fixed.acidity, pH)) + 
  geom_jitter(alpha=0.2) + 
  stat_smooth(method = "lm", formula = y ~ x)

```


We see fixed acidity increase as pH decreases.


```{r echo=FALSE}
ggplot(df, aes(fixed.acidity, density)) + 
  geom_point(alpha=0.2) +  
  stat_smooth(method = "lm", formula = y ~ x)

```

Fixed acidity increases as density increases.

```{r echo=FALSE}
ggplot(df, aes(quality, alcohol)) + 
  geom_jitter(alpha=0.2) +  
  stat_smooth(method = "lm", formula = y ~ x)

```

Quality seem to be incresing with higher alhocol.

# Bivariate Analysis

> Alcohol has the highest positive correlation with quality, followed by sulphates, citric.acid, and fixed.acidity. We explored this further by comparing alcohol and density. Density rises as alcohol drops. Given their negative and positive correlations, this is to be expected. It's the same with fixed acidity and ph, which has the highest negative correlation coefficient among our fields. Density and fixed acidity correlate positively, and trend in the same direction. More density means more fixed acidity.


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE}
ggplot(df, aes(y = fixed.acidity, x = pH, color = quality.factor)) + 
        geom_point()
```

The graph shows that the higher the acidity the lower pH. We can see any relationships related to color

```{r echo=FALSE}
ggplot(df, aes(y = fixed.acidity, x = density, color = quality.factor)) + 
        geom_jitter(alpha=0.7)
```

There seems to be a strong direct relationship where density increases as acidity increases. 
Adding quality by color shows lower quality (4 and 5) are high density and low acidity compared to (6, 7, 8).

```{r echo=FALSE}
ggplot(df, aes(y = residual.sugar, x = alcohol, color = quality.factor)) + 
        geom_jitter()
```

Comparing the residual sugar and alcohol shows most of the red wine is below 4 which means they 
would be off-dry wines and range from 9-13% in alcohol content.


```{r echo=FALSE}
ggplot(df, aes(y = volatile.acidity, x = pH, color = alcohol_content)) + 
        geom_jitter() + 
        
        # creates multiple graphs for each quality 
        facet_wrap(~quality.factor, ncol = 3)
```

Each graph is colored by alcohol_content and seperated by quality. We see medium and high alcohol 
content more prominent in higher quality wines


# Multivariate Analysis

There seems to be a relationship between quality and alcohol content. Where higher quality wine 
usually contain more alcohol

### Were there any interesting or surprising interactions between features?
I would not say that was surprised by my finding as I would expect higher alcohol content in better wines

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}
ggplot(df, aes(alcohol, fill = "#000")) + 
        geom_histogram(binwidth = .05) +
        scale_x_continuous(breaks = 8:15) +
        
        # adding X, Y, and graph titles 
        labs(title = "Red Wine Alcohol Percentage", 
             x = "Alcohol Percentage",
             y = "Number of Wines", size = 12) + 
        
        # changing the size of the titles
        theme(plot.title = element_text(size = rel(2)), 
              axis.title.y = element_text(size = rel(1.5), angle = 90), 
              axis.title.x = element_text(size = rel(1.5), angle = 0)) + 
        
        # removes the legend since it is not required
        guides(fill = F)
```

### Description One
This graph shows the greatest alcohol % to be around 9.5% and we see a gradual decrease as the % increases 
or decreases. This is probably becuase that's the optimal alcohol % that is time efficient 
(as in the amount of time needed for fermentation) for wholesale, or the amount most consumers prefer.

### Plot Two
```{r echo=FALSE, Plot_Two}
ggplot(df, aes(y = volatile.acidity, x = pH, color = alcohol_content)) + 
        geom_jitter() + 
        facet_wrap(~quality, ncol = 2) + 
        scale_color_brewer(palette = 'Set2') +
        labs(color = "Alcohol Content") +
        
        labs(title = "Volatile Acidity of pH by Alcohol Content", 
             y = "Volatile Acidity", 
             x = "pH", 
             size = 12) + 
        
        theme(plot.title = element_text(size = rel(2)), 
              axis.title.y = element_text(size = rel(1.5), angle = 90), 
              axis.title.x = element_text(size = rel(1), angle = 0))

```

### Description Two

This plot shows how the alcohol content changes with quality, mostly high content wines 
are promenent in quality 8 while the opposite is true for quality 3. There is also a small 
down shift towards volatile acidity as quality increases, I suspect to lower the vinger taste.

### Plot Three

```{r echo=FALSE, Plot_Three}
ggplot(df, aes(y = fixed.acidity, x = density, color = quality.factor)) + 
        geom_point() +
        scale_color_brewer(palette = 'OrRd') +
        labs(color="Quality")
        labs(title = "Fixed Acidity of density by quality", 
             x = "Density", 
             y = "Fixed Acidity", 
             size = 12) + 
        
        theme(plot.title = element_text(size = rel(2)), 
              axis.title.y = element_text(size = rel(1), angle = 90), 
              axis.title.x = element_text(size = rel(1.5), angle = 0))
```

### Description Three
The graph shows how acidity increases as density increases. It is also able to determine loosely, how the higher quality is more acidic.

------

# Reflection

I found exploring a red wines dataset very interesting while I had a suspicion than 
higher quality wines generally have higher alcohol content it was a fun to explore and 
actually confirm that assumption. 
I found challenging to actually use R studio and R language, in general, 
I'm used to programming in Python so this was quite a challenge but at the end quite rewarding 
to know another programing language even on a basic level.
For future, I could perhaps work out a machine learning model that would predict a price or 
quality of wine based on data provided
